Generalized Bradley-Terry models

### 2025-02-17 inst/jss_hyper3_real.Rnw
more work

\begin{figure}[htbp]
  \begin{center}
<<plotproflikeoldfirm, echo=FALSE, fig=TRUE>>=
plot(lam,logLike - max(logLike),type='b')
abline(v=1,col='gray')
segments(x0=1.785,y0=0.1,y1=-1,col='red')
abline(h = c(0,-2),col='gray')
@
\caption{A\label{profilesupportlambda} profile support function for the
  home-team advantage $\lambda$.  Vertical gray line shows
  $H_0\colon\lambda =1$, horizontal lines show the 
  two-units-of-support limits; small red bar shows
  $\hat{\lambda}\simeq 1.785$}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
  \begin{center}
<<makecontour, fig=TRUE, cache=TRUE>>=
n <- 2
lambda <- seq(from=1, to=3, len=n)
D <- seq(from=0.1, to=0.6, len=n)
V <- as.matrix(expand.grid(lambda, D))
LL <- apply(V, 1, f)
LL <- LL - max(LL)
contour(lambda, D, matrix(LL, n, n))
@
\caption{Contours of support\label{contourlambdaD} for a grid}
\end{center}
\end{figure}

Thus \code{f()} gives us a likelihood function for
$\mathbf{v}=(\lambda,D)$.  We may apply \code{f()} over a grid and
thereby determine a likelihood surface, shown in
figure~\ref{contourlambdaD}.  Further, we may find the evaluate using
\code{optim()}:

@@

<<testlambdazero, cache=TRUE>>=
jj <- optimize(f=function(D){f(c(lambda=1,D))}, interval=c(0.1, 0.4), maximum=TRUE)
jj
@

>>>>

\begin{figure}[htbp]
  \begin{center}
<<plotproflikeoldfirm, echo=FALSE>>=
plot(lam,logLike - max(logLike),type='b')
abline(v=1,col='gray')
segments(x0=1.785,y0=0.1,y1=-1,col='red')
abline(h = c(0,-2),col='gray')
@
\caption{A\label{profilesupportlambda} profile support function for the
  home-team advantage $\lambda$.  Vertical gray line shows
  $H_0\colon\lambda =1$, horizontal lines show the 
  two-units-of-support limits; small red bar shows
  $\hat{\lambda}\simeq 1.785$}
\end{center}
\end{figure}

@@

Thus \code{f()} gives us a likelihood function for
$\mathbf{v}=(\lambda,D)$.  We may find the evaluate using
\code{optim()}:

@@

Then

<<showevaluate, cache=TRUE>>=
maxp(home_draw_away3(RCLF3_table, lambda = maxv[1], D=maxv[2]))
@

Such estimates have rich predictive power.  For example, if Rangers
were to play Celtic with both teams away [thus nullifying the home
  ground advantage: $\lambda=1$ in Table~\ref{windrawloss}], the
probability of win/draw/loss for Rangers would be
$\left(\frac{p_R}{(1+D)(p_R+p_C)};\frac{D}{1+D};\frac{p_R}{(1+D)(p_R+p_C)}\right)$,
estimated at $(0.39;0.27;0.34)$.

\begin{figure}[htbp]
  \begin{center}
<<makecontour, cache=TRUE>>=
n <- 2
lambda <- seq(from=1, to=3, len=n)
D <- seq(from=0.1, to=0.6, len=n)
V <- as.matrix(expand.grid(lambda, D))
LL <- apply(V, 1, f)
LL <- pmax(-20, LL - max(LL))
contour(lambda, D, matrix(LL, n, n),levels = -2*(0:9), xlab=expression(lambda),ylab=expression(D))
points(maxv[1], maxv[2], pch=16, col='red')
@
\caption{Contours of support\label{contourlambdaD} evaluated over
  a grid.  Maximum likelihood estimate for $(\lambda, D)$ shown as
  a red dot at the evaluate}
  \end{center}
\end{figure}

<<testlambdazero, cache=TRUE, include=FALSE>>=
jj <- optimize(f=function(D){f(c(lambda=1,D))}, interval=c(0.1, 0.4), maximum=TRUE)
@

Further, we may calculate a support function over the
$(\lambda,D)$-plane, shown in figure~\ref{contourlambdaD}.  For
example, we may test the hypothesis that $\lambda=1$ [corresponding to
  the right hand side of figure~\ref{contourlambdaD}] for which the
support is about \Sexpr{round(f(maxv) - jj$objective,2)}, far
exceeding Edwards's two units of support per degree of freedom
criteron.  Alternatively, we can observe that this lies in the tail
region of its asymptotic null $\chi^2_4$ distribution with a $p$-value
of about \Sexpr{signif(pchisq(f(maxv) -
  jj$objective,lower.tail=FALSE,df=2),digits=2)}.
